A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data

### 2024-01-11 tidyindex.qmd
modify Nicolas' affiliation

---
title: "A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data"
format:
  tandf-pdf:
    keep-tex: true  
pdf-engine: pdflatex
author:
  - name: H. Sherry Zhang
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, Victoria
        country: Australia
    orcid: 0000-0002-7122-1463
    email: huize.zhang@monash.edu
  - name: Dianne Cook
    affiliation: 
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, Victoria
        country: Australia
    orcid: 0000-0002-3813-7155
  - name: Ursula Laa
    affiliation: 
      - name: University of Natural Resources and Life Sciences
        department: Institute of Statistics
        address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
        city: Vienna
        country: Austria
    orcid: 0000-0002-0249-6439
  - name: Nicolas Langrené
    affiliation: 
      - name: BNU-HKBU United International College
        department: Department of Mathematical Sciences
        address: 2000 Jintong Road, Tangjiawan, Zhuhai
        city: Zhuhai, Guangdong
        country: China
    orcid: 0000-0001-7601-4618
  - name: Patricia Menéndez
    affiliation:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, Victoria
        country: Australia
    orcid: 0000-0003-0701-6315
abstract: |
  Indexes are useful for summarizing multivariate information into single metrics for monitoring, communicating, and decision-making. While most work has focused on defining new indexes for specific purposes, more attention needs to be directed towards making it possible to understand index behavior in different data conditions, and to determine how their structure affects their values and variation in values. Here we discuss a modular data pipeline recommendation to assemble indexes. It is universally applicable to index computation and allows investigation of index behavior as part of the development procedure. One can compute indexes with different parameter choices, adjust steps in the index definition by adding, removing, and swapping them to experiment with various index designs, calculate uncertainty measures, and assess indexes' robustness. The paper presents three examples to illustrate the pipeline framework usage: comparison of two different indexes designed to monitor the spatio-temporal distribution of drought in Queensland, Australia; the effect of dimension reduction choices on the Global Gender Gap Index (GGGI) on countries' ranking; and how to calculate bootstrap confidence intervals for the Standardized Precipitation Index (SPI). The methods are supported by a new R package, called `tidyindex`.
keywords:
  - indexes
  - data pipeline
  - software design
  - uncertainty
  - decision-making
bibliography: bibliography.bib  
editor_options: 
  chunk_output_type: console
---

@@

This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship. The article is created using Quarto [@Allaire_Quarto_2022] in R [@R]. The source code for reproducing this paper can be found at: <https://github.com/huizezhang-sherry/paper-tidyindex>.

>>>>

---
title: "A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data"
format:
  tandf-pdf:
    keep-tex: true  
    linestretch: 2
pdf-engine: pdflatex
author:
  - name: H. Sherry Zhang
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, Victoria
        country: Australia
    orcid: 0000-0002-7122-1463
    email: huize.zhang@monash.edu
  - name: Dianne Cook
    affiliation: 
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, Victoria
        country: Australia
    orcid: 0000-0002-3813-7155
  - name: Ursula Laa
    affiliation: 
      - name: University of Natural Resources and Life Sciences
        department: Institute of Statistics
        address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
        city: Vienna
        country: Austria
    orcid: 0000-0002-0249-6439
  - name: Nicolas Langrené
    affiliation: 
      - name: BNU-HKBU United International College
        department: Guangdong Provincial Key Laboratory of Interdisciplinary Research and Application for Data Science
        address: 2000 Jintong Road, Tangjiawan, Zhuhai
        city: Zhuhai, Guangdong
        country: China
    orcid: 0000-0001-7601-4618
  - name: Patricia Menéndez
    affiliation:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, Victoria
        country: Australia
    orcid: 0000-0003-0701-6315
abstract: |
  Indexes are useful for summarizing multivariate information into single metrics for monitoring, communicating, and decision-making. While most work has focused on defining new indexes for specific purposes, more attention needs to be directed towards making it possible to understand index behavior in different data conditions, and to determine how their structure affects their values and variation in values. Here we discuss a modular data pipeline recommendation to assemble indexes. It is universally applicable to index computation and allows investigation of index behavior as part of the development procedure. One can compute indexes with different parameter choices, adjust steps in the index definition by adding, removing, and swapping them to experiment with various index designs, calculate uncertainty measures, and assess indexes' robustness. The paper presents three examples to illustrate the pipeline framework usage: comparison of two different indexes designed to monitor the spatio-temporal distribution of drought in Queensland, Australia; the effect of dimension reduction choices on the Global Gender Gap Index (GGGI) on countries' ranking; and how to calculate bootstrap confidence intervals for the Standardized Precipitation Index (SPI). The methods are supported by a new R package, called `tidyindex`.
keywords:
  - indexes
  - data pipeline
  - software design
  - uncertainty
  - decision-making
bibliography: bibliography.bib  
editor_options: 
  chunk_output_type: console
---

@@

This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship. Nicolas Langrené acknowledges the partial support of the Guangdong Provincial Key Laboratory IRADS (2022B1212010006, R0400001-22) and the UIC Start-up Research Fund UICR0700041-22. The article is created using Quarto [@Allaire_Quarto_2022] in R [@R]. The source code for reproducing this paper can be found at: <https://github.com/huizezhang-sherry/paper-tidyindex>.

### 2024-01-11 tidyindex.tex
modify Nicolas' affiliation

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Econometrics and Business
Statistics, Monash University, Melbourne,
Victoria, Australia\\ \textsuperscript{2} Institute of
Statistics, University of Natural Resources and Life
Sciences, Vienna, Austria\\ \textsuperscript{3} Department of
Mathematical Sciences, BNU-HKBU United International College, Zhuhai,
Guangdong, China
\begin{abstract}
Indexes are useful for summarizing multivariate information into single
metrics for monitoring, communicating, and decision-making. While most
work has focused on defining new indexes for specific purposes, more
attention needs to be directed towards making it possible to understand
index behavior in different data conditions, and to determine how their
structure affects their values and variation in values. Here we discuss
a modular data pipeline recommendation to assemble indexes. It is
universally applicable to index computation and allows investigation of
index behavior as part of the development procedure. One can compute
indexes with different parameter choices, adjust steps in the index
definition by adding, removing, and swapping them to experiment with
various index designs, calculate uncertainty measures, and assess
indexes' robustness. The paper presents three examples to illustrate the
pipeline framework usage: comparison of two different indexes designed
to monitor the spatio-temporal distribution of drought in Queensland,
Australia; the effect of dimension reduction choices on the Global
Gender Gap Index (GGGI) on countries' ranking; and how to calculate
bootstrap confidence intervals for the Standardized Precipitation Index
(SPI). The methods are supported by a new R package, called
\texttt{tidyindex}.
\end{abstract}
\begin{keywords}
\def\sep{;\ }
indexes\sep data pipeline\sep software design\sep uncertainty\sep 
decision-making
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, frame hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, enhanced]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

This work is funded by a Commonwealth Scientific and Industrial Research
Organisation (CSIRO) Data61 Scholarship. The article is created using
Quarto (Allaire et al. 2022) in R (R Core Team 2021). The source code
for reproducing this paper can be found at:
\url{https://github.com/huizezhang-sherry/paper-tidyindex}.

>>>>

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Econometrics and Business
Statistics, Monash University, Melbourne,
Victoria, Australia\\ \textsuperscript{2} Institute of
Statistics, University of Natural Resources and Life
Sciences, Vienna, Austria\\ \textsuperscript{3} Guangdong Provincial Key
Laboratory of Interdisciplinary Research and Application for Data
Science, BNU-HKBU United International College, Zhuhai, Guangdong, China
\begin{abstract}
Indexes are useful for summarizing multivariate information into single
metrics for monitoring, communicating, and decision-making. While most
work has focused on defining new indexes for specific purposes, more
attention needs to be directed towards making it possible to understand
index behavior in different data conditions, and to determine how their
structure affects their values and variation in values. Here we discuss
a modular data pipeline recommendation to assemble indexes. It is
universally applicable to index computation and allows investigation of
index behavior as part of the development procedure. One can compute
indexes with different parameter choices, adjust steps in the index
definition by adding, removing, and swapping them to experiment with
various index designs, calculate uncertainty measures, and assess
indexes' robustness. The paper presents three examples to illustrate the
pipeline framework usage: comparison of two different indexes designed
to monitor the spatio-temporal distribution of drought in Queensland,
Australia; the effect of dimension reduction choices on the Global
Gender Gap Index (GGGI) on countries' ranking; and how to calculate
bootstrap confidence intervals for the Standardized Precipitation Index
(SPI). The methods are supported by a new R package, called
\texttt{tidyindex}.
\end{abstract}
\begin{keywords}
\def\sep{;\ }
indexes\sep data pipeline\sep software design\sep uncertainty\sep 
decision-making
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, enhanced, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi

\setstretch{2}
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

This work is funded by a Commonwealth Scientific and Industrial Research
Organisation (CSIRO) Data61 Scholarship. Nicolas Langrené acknowledges
the partial support of the Guangdong Provincial Key Laboratory IRADS
(2022B1212010006, R0400001-22) and the UIC Start-up Research Fund
UICR0700041-22. The article is created using Quarto (Allaire et al.
2022) in R (R Core Team 2021). The source code for reproducing this
paper can be found at:
\url{https://github.com/huizezhang-sherry/paper-tidyindex}.
